---
title: "covid 19"
author: "Michelle Cervantes"
date: "10/19/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 
### This is contained in a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
# Question 1 - Is dying independent of sex




Table of patients:
```{r}

# cross tabulation, patient state by gender

with(kRdata, table(Sex, result))

```

## We create two bar charts - one based on frequency and the other on percent

```{r}
barchartGC(~result + Sex,data=kRdata, main ="Frequency Bar Chart")
```

  
### this bar chart is displaying in total frequency terms our data.  Clearly there are a lot less dead people than alive people as is reflected by the sizes of the bars in the display.  

```{r}
barchartGC(~result + Sex,data=kRdata, type="Percent", main = "Frequency Bar Chart")
```

Based off this chart is hows that out of 100% in the 'alive' group, about 59% of them are women and about 41% are men. In the 'dead' group, about 40% are women and about 62% are men.  

## Numerical results

```{r}
table1 <- xtabs(~result + Sex,data=kRdata)
rowPerc(table1)
colPerc(table1)
```

Of the people who are alive, 56% of them are female and 44% of them are male and of the people who are dead, 36% of them are female and 64% of them are male. It looks like males are dying a lot more frequently than females are.


## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

On the Fisher's exact test chart, the odds ratio tells us that the odds ratio of 2.3 : 1 means that the odds of a man dying is 2.3 chances higher than a woman. Our p-value is 0.0009, which tells us that the probability of data occurring again is 9 out of 10,000 which is not very likely but it can be possible since this disease is more harmful to men than women. 

## Conclusions

To summarize the results, we noticed how COVID19 is far more fatal to men than women, so fatal that the chances of a man dying to covid is 2.3 chances higher than a woman. To answer the question, "Is dying independent of sex?" I reject the null hypothesis and would say that no. In this case, dying seems to be dependent on sex because of how frequently men are dying in comparison to women. 
